The Multi-Stencil Language

### 2016-01-29 ARTICLES/IJPP/article/compiler.tex
article IJPP language corrections Helene 27/01/2016
article IJPP corrections Helene 29/01/2016

\begin{algorithm}
\caption{MSL Compiler}
\label{alg:compiler}
\begin{algorithmic}[1]
\Procedure{compileMSL} {file}
\\$\Gamma$ = parser(file)
\\$\Gamma_{sync} = F_{sync}(\Gamma,0)$
\\$\Gamma_{dep} = F_{dep}(\Gamma_{sync},0)$
\\removeNSH($\Gamma_{dep}$)
\\$TSP = BDT(L^{-1}(\Gamma_{dep}))$
\\$TSP = F_{tsp}(TSP,root(TSP))$
\If {data parallel}
\\$TSP = F_{tsp}(TSP)$
\EndIf
\\dump(TSP)
\EndProcedure
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{$F_{tsp}$}
\label{alg:tsp}
\begin{algorithmic}[1]
\Procedure{$F_{tsp}$} {$TSP$,root,succ}
\For {$s \in succ$}
\If {type($s$)=$S$ or type($s$)=$P$}
\\$F_{tsp}(TSP,s,successors(s))$
\If {type($s$)=type(root)}
\\$TSP = (V_{tsp} \setminus s,E_{tsp} \cup replace(E_{rooted(s)},s,root))$
\EndIf
\EndIf
\EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}

Using MSL, it is possible to ask for data parallelization of the application, or for an hybrid parallelization. Even though the MSL language is not dedicated to produce very optimized stencil codes, but to produces the parallel pattern of the application, building the $TSP$ tree make available an easy optimization when the data parallelization technique is the only one used. This optimization consists in proposing a valid merge of some computation kernels inside a single space loop. As a result, the user can use this valid fusion of kernels or not when implementing those. The fusion function $F_{fus}$ is described in Algorithm~\ref{alg:fus}, where $parent(k)$ returns the parent vertex of $k$ in the tree, and where $k_{i;j}$ represents the fusion of $k_i$ and $k_j$ keeping the sequential order $i;j$ if $i$ is computed before $j$ in $TSP$.

\begin{algorithm}
\caption{$F_{fus}$}
\label{alg:fus}
\begin{algorithmic}[1]
\Procedure{$F_{fus}$} {$TSP(V,E)$}
\For {$(k_i,k_j) \in V^2$}
\If {$type(k_i)=type(k_j)=comp$}
\If {parent($k_i$)==parent($k_j$)}
\If {parent($k_i$)==$S$}
\If {$D_i==D_j$}
\\propose the fusion $k_{i;j}$
\EndIf
\EndIf
\If {parent($k_i$)==$P$}
\If {$D_i==D_j$ and $R_i \cap R_j \neq \emptyset$}
\\propose the fusion $k_{i;j}$
\EndIf
\EndIf
\EndIf
\EndIf
\EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}

We are not arguing that such a simple fusion algorithm could be as good as complex cache optimization techniques which can be found in stencil DSLs for example~\cite{spaaTangCKLL11}. However, this fusion takes place at a different level and can bring performance improvments as it will be illustrated in Section~\ref{sect:eval}. This fusion algorithm relies on very simple statements:
\begin{itemize}
\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are, by construction, data dependant. As a result, what is written by the first one is read by the second one. Thus, at least one data is common to those computations (the one written by $k_i$). Thus, if the computation domains verify $D_i=D_j$, the fusion of $k_i$ and $k_j$ will decrease cache misses.
\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $P$ in TSP are not, by construction, data dependant. However, if the computation domains verify $D_i=D_j$, and if $R_i \cap R_j \neq \emptyset$ cache misses could also be decreased by the fusion $k_{i;j}$.
\end{itemize}

\todo{solve order problem in $\Gamma_{dep}$ and $TSP$}

@@

@@

%====BIG EXAMPLE
\begin{figure}[ht!]
\begin{center}
% INITIAL GRAPH
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (c0) at (0,0) {$c_0$};
   \node[] (star1) at (1,0) {$*_1$};
   \node[] (c1) at (2,0) {$c_1^*$};
   \node[] (c2) at (3,0.5) {$c_2$};
   \node[] (star4) at (3,1.5) {$*_4$};
   \node[] (c3) at (3,-0.5) {$c_3$};
   \node[] (c4) at (4,0.5) {$c_4^*$};
   \node[] (c5) at (4,-0.5) {$c_5$};
   \node[] (c6) at (5,0.5) {$c_6$};
   \node[] (c7) at (6,0) {$c_7$};
   \node[] (star8) at (7,0) {$*_8$};
   \node[] (c8) at (8,0) {$c_8^*$};

@@

  \path[->]
    (c0s) edge node [above] {$c_0$} (c0d)
    (c0d) edge node [above] {$*_1$} (star1d)
    (star1d) edge node [above] {$c_1^*$} (c1d)
    (c1d) edge node [above] {$c_2$} (c2d)
          edge node [above] {$c_3$} (c3d)
          edge [bend left=80] node [above] {$*_4$} (c2d)
    (c2d) edge node [above] {$c_4^*$} (c4d)
    (c4d) edge node [above] {$c_6$} (c6d)
    (c6d) edge node [above] {$c_7$} (c7d)
    (c7d) edge node [above] {$c_8$} (star8d)
    (star8d) edge node [above] {$c_8$} (c8d)
    (c3d) edge node [above] {$c_5$} (c6d);
  \end{tikzpicture}
}
\\
% TRANSFORMATIONS
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (4,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (4,-0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (5,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

   %reduction c0 c1
   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,2.5) {$c_0$};
   \node[] (star1) at (1.5,2.5) {$*_1$};
   \node[] (c1) at (2,1.5) {$c_1^*$};

   %reduction c7 c8
   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
   \node[] (c7) at (6.5,2.5) {$c_7$};
   \node[] (star8) at (7.5,2.5) {$*_8$};
   \node[] (c8) at (8,1.5) {$c_8$};

  \path[->]
    (c0s) edge node {} (c1d)
    (c1d) edge node [above] {$c_2$} (c2d)
          edge node [above] {$c_3$} (c3d)
          edge [bend left=80] node [above] {$*_4$} (c2d)
    (c2d) edge node [above] {$c_4^*$} (c4d)
    (c4d) edge node [above] {$c_6$} (c6d)
    (c6d) edge node [above] {} (c8d)
    (c3d) edge node [above] {$c_5$} (c6d)

@@

   %reduction c0 c1
   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,2.5) {$c_0$};
   \node[] (star1) at (1.5,2.5) {$*_1$};
   \node[] (c1) at (2,1.5) {$c_1^*$};

   %reduction c7 c8
   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
   \node[] (c7) at (6.5,2.5) {$c_7$};
   \node[] (star8) at (7.5,2.5) {$*_8$};
   \node[] (c8) at (8,1.5) {$c_8$};

   %reduction c3 c5
   \node[] (s5) at (4.5,-2) {$\mathcal{S}$};
   \node[] (c3) at (4,-1) {$c_3$};
   \node[] (c5) at (5,-1) {$c_5$};

   %reduction c2 *4
   \node[] (p0) at (3.5,1) {$\mathcal{P}$};
   \node[] (c2) at (3,2) {$c_2$};
   \node[] (star4) at (4,2) {$*_4$};

  \path[->]
    (c0s) edge node {} (c1d)
    (c1d) edge node [above] {} (c2d)
    (c2d) edge node [above] {$c_4^*$} (c4d)
    (c4d) edge node [above] {$c_6$} (c6d)
    (c6d) edge node [above] {} (c8d)
    (c1d) edge [bend right] node [above] {} (c6d)

@@

    %reduction c2 *4
    (p0) edge node {} (c2)
         edge node {} (star4);
  \end{tikzpicture}
}
\end{center}
\caption{Big example}
\label{fig:example}
\end{figure}

\begin{figure}[ht!]
\ContinuedFloat
\begin{center}
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

   %reduction c0 c1
   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,2.5) {$c_0$};
   \node[] (star1) at (1.5,2.5) {$*_1$};
   \node[] (c1) at (2,1.5) {$c_1^*$};

   %reduction c7 c8
   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
   \node[] (c7) at (6.5,2.5) {$c_7$};
   \node[] (star8) at (7.5,2.5) {$*_8$};
   \node[] (c8) at (8,1.5) {$c_8$};

   %reduction c3 c5
   \node[] (s5) at (4.5,-2) {$\mathcal{S}$};
   \node[] (c3) at (4,-1) {$c_3$};
   \node[] (c5) at (5,-1) {$c_5$};

   %reduction c2 c6
   \node[] (s6) at (4.5,1) {$\mathcal{S}$};
   \node[] (p0) at (3.5,2) {$\mathcal{P}$};
   \node[] (c2) at (3,3) {$c_2$};
   \node[] (star4) at (4,3) {$*_4$};
   \node[] (s7) at (5.5,2) {$\mathcal{S}$};
   \node[] (c4) at (5,3) {$c_4^*$};
   \node[] (c6) at (6,3) {$c_6$};

@@

   %reduction c0 c1
   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,2.5) {$c_0$};
   \node[] (star1) at (1.5,2.5) {$*_1$};
   \node[] (c1) at (2,1.5) {$c_1^*$};

   %reduction c7 c8
   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
   \node[] (c7) at (6.5,2.5) {$c_7$};
   \node[] (star8) at (7.5,2.5) {$*_8$};
   \node[] (c8) at (8,1.5) {$c_8$};

   \node[] (p1) at (4.5,0.5) {$\mathcal{P}$};
   %reduction c3 c5
   \node[] (s5) at (5.5,1.5) {$\mathcal{S}$};
   \node[] (c3) at (5,2.5) {$c_3$};
   \node[] (c5) at (6,2.5) {$c_5$};
   %reduction c2 c6
   \node[] (s6) at (3.5,1.5) {$\mathcal{S}$};
   \node[] (p0) at (2.8,2.5) {$\mathcal{P}$};
   \node[] (c2) at (2.3,3.5) {$c_2$};
   \node[] (star4) at (3.3,3.5) {$*_4$};
   \node[] (s7) at (4.2,2.5) {$\mathcal{S}$};
   \node[] (c4) at (3.8,3.5) {$c_4^*$};
   \node[] (c6) at (4.8,3.5) {$c_6$};

@@

   %reduction c0 c1
   \node[] (s0) at (1.5,2.5) {$\mathcal{S}$};
   \node[] (s1) at (1,3.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,4.5) {$c_0$};
   \node[] (star1) at (1.5,4.5) {$*_1$};
   \node[] (c1) at (2,3.5) {$c_1^*$};

   %reduction c7 c8
   \node[] (s2) at (7.5,1.5) {$\mathcal{S}$};
   \node[] (s3) at (7,2.5) {$\mathcal{S}$};
   \node[] (c7) at (6.5,3.5) {$c_7$};
   \node[] (star8) at (7.5,3.5) {$*_8$};
   \node[] (c8) at (8,2.5) {$c_8$};

   \node[] (p1) at (4.5,2.5) {$\mathcal{P}$};
   %reduction c3 c5
   \node[] (s5) at (5.5,3.5) {$\mathcal{S}$};
   \node[] (c3) at (5,4.5) {$c_3$};
   \node[] (c5) at (6,4.5) {$c_5$};
   %reduction c2 c6
   \node[] (s6) at (3.5,3.5) {$\mathcal{S}$};
   \node[] (p0) at (2.8,4.5) {$\mathcal{P}$};
   \node[] (c2) at (2.3,5.5) {$c_2$};
   \node[] (star4) at (3.3,5.5) {$*_4$};
   \node[] (s7) at (4.2,4.5) {$\mathcal{S}$};
   \node[] (c4) at (3.8,5.5) {$c_4^*$};
   \node[] (c6) at (4.8,5.5) {$c_6$};

@@

    (p1) edge node {} (s6)
         edge node {} (s5)
    %reduction c3 c5
    (s5) edge node {} (c3)
         edge node {} (c5)
    %reduction c2 c6
    (s6) edge node {} (p0)
         edge node {} (s7)
    %reduction c2 *4
    (p0) edge node {} (c2)
         edge node {} (star4)
    (s7) edge node {} (c4)
         edge node {} (c6);
  \end{tikzpicture}
}
\end{center}
\caption{Big example}
\label{fig:example}
\end{figure}

>>>>

\begin{algorithm}
\caption{MSL Compiler}
\label{alg:compiler}
\begin{algorithmic}[1]
\Procedure{compileMSL} {file}
\State $\Gamma$ = parser(file)
\State $\Gamma_{sync} = F_{sync}(\Gamma,0)$
\State $\Gamma_{dep} = F_{dep}(\Gamma_{sync},0)$
\State removeNSH($\Gamma_{dep}$)
\State $TSP = BDT(L^{-1}(\Gamma_{dep}))$
\State $TSP = F_{tsp}(TSP,root(TSP))$
\If {data parallel}
\State $TSP = F_{tsp}(TSP)$
\EndIf
\State dump(TSP)
\EndProcedure
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{$F_{tsp}$}
\label{alg:tsp}
\begin{algorithmic}[1]
\Procedure{$F_{tsp}$} {$TSP$,root,succ}
\For {$s \in succ$}
\If {type($s$)=$S$ or type($s$)=$P$}
\State $F_{tsp}(TSP,s,successors(s))$
\If {type($s$)=type(root)}
\State $V_{tsp} = V_{tsp} \setminus s$
\State $E_{tsp} = E_{tsp} \cup replace(E_{rooted(s)},s,root)$
\State $TSP = (V_{tsp},E_{tsp})$
\EndIf
\EndIf
\EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}

Using MSL, it is possible to ask for data parallelization of the application, or for an hybrid parallelization. Even though the MSL language is not dedicated to produce very optimized stencil codes, but to produces the parallel pattern of the application, building the $TSP$ tree make available an easy optimization when the data parallelization technique is the only one used. This optimization consists in proposing a valid merge of some computation kernels inside a single space loop. As a result, the user can use this valid fusion of kernels or not when implementing those. The fusion function $F_{fus}$ is described in Algorithm~\ref{alg:fus}, where $parent(k)$ returns the parent vertex of $k$ in the tree, and where $k_{i;j}^{fus}$ represents the fusion of $k_i$ and $k_j$ keeping the sequential order $i;j$ if $i$ is computed before $j$ in $TSP$.

\begin{algorithm}
\caption{$F_{fus}$}
\label{alg:fus}
\begin{algorithmic}[1]
\Procedure{$F_{fus}$} {$TSP(V,E)$}
\For {$(k_i,k_j) \in V^2$}
\If {parent($k_i$)==parent($k_j$)}
\If {$type(k_i)=type(k_j)=comp$}
\If {parent($k_i$)==$S$}
\If {$D_i==D_j$}
\State propose the fusion $k_{i;j}^{fus}$
\EndIf
\ElsIf {parent($k_i$)==$P$}
\If {$D_i==D_j$ and $R_i \cap R_j \neq \emptyset$}
\State propose the fusion $k_{i;j}^{fus}$
\EndIf
\EndIf
\EndIf
\EndIf
\EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}

We are not arguing that such a simple fusion algorithm could be as good as complex cache optimization techniques which can be found in stencil DSLs for example~\cite{spaaTangCKLL11}. However, this fusion takes place at a different level and can bring performance improvments as it will be illustrated in Section~\ref{sect:eval}. This fusion algorithm relies on very simple statements:
\begin{itemize}
\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are, by construction, data dependant. As a result, what is written by the first one is read by the second one. Thus, at least one data is common to those computations (the one written by $k_i$). Thus, if the computation domains verify $D_i=D_j$, the fusion of $k_i$ and $k_j$ will decrease cache misses.
\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $P$ in TSP are not, by construction, data dependant. However, if the computation domains verify $D_i=D_j$, and if $R_i \cap R_j \neq \emptyset$ cache misses could also be decreased by the fusion $k_{i;j}^{fus}$.
\end{itemize}

@@

MSL code~\ref{fig:exmsl}. 
\begin{equation*}
\Gamma = [k_0,k_1,k_2,k_3,k_4,k_0,k_6,k_7,k_8]
\end{equation*}

\begin{equation*}
\Gamma_{sync} = [k_0,k_{0;1}^{sync},k_1,k_2,k_3,k_{1;4}^{sync},k_4,k_0,k_6,k_7,k_{7;8}^{sync},k_8]
\end{equation*}

where $k_{0;1}^{sync}$...

$\Gamma_{dep}$~\ref{} and $L^{-1}(\Gamma_{dep})$~\ref{}

@@

%====BIG EXAMPLE
\begin{figure}[ht!]
\begin{center}
% INITIAL GRAPH
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (c0) at (0,0) {$k_0$};
   \node[] (star1) at (1,0) {$k_{0;1}^{sync}$};
   \node[] (c1) at (2,0) {$k_1$};
   \node[] (c2) at (3,0.5) {$k_2$};
   \node[] (star4) at (3,1.5) {$k_{1;4}^{sync}$};
   \node[] (c3) at (3,-0.5) {$k_3$};
   \node[] (c4) at (4,0.5) {$k_4$};
   \node[] (c5) at (4,-0.5) {$k_5$};
   \node[] (c6) at (5,0.5) {$k_6$};
   \node[] (c7) at (6,0) {$k_7$};
   \node[] (star8) at (7,0) {$k_{7;8}^{sync}$};
   \node[] (c8) at (8,0) {$k_8$};

@@

  \path[->]
    (c0s) edge node [above] {$k_0$} (c0d)
    (c0d) edge node [above] {$k_{0;1}^{sync}$} (star1d)
    (star1d) edge node [above] {$k_1$} (c1d)
    (c1d) edge node [above] {$k_2$} (c2d)
          edge node [above] {$k_3$} (c3d)
          edge [bend left=80] node [above] {$k_{1;4}^{sync}$} (c2d)
    (c2d) edge node [above] {$k_4$} (c4d)
    (c4d) edge node [above] {$k_6$} (c6d)
    (c6d) edge node [above] {$k_7$} (c7d)
    (c7d) edge node [above] {$k_{7;8}^{sync}$} (star8d)
    (star8d) edge node [above] {$k_8$} (c8d)
    (c3d) edge node [above] {$k_0$} (c6d);
  \end{tikzpicture}
}
\end{center}
\caption{$\Gamma_{dep}$ and $L^{-1}(\Gamma_{dep})$}
\label{fig:dep}
\end{figure}

\begin{figure}[ht!]
\begin{center}
% TRANSFORMATIONS
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (4,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (4,-0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (5,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

   %reduction c0 c1
   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,2.5) {$k_0$};
   \node[] (star1) at (1.5,2.5) {$k_{0;1}^{sync}$};
   \node[] (c1) at (2,1.5) {$k_1$};

   %reduction c7 c8
   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
   \node[] (c7) at (6.5,2.5) {$k_7$};
   \node[] (star8) at (7.5,2.5) {$k_{7;8}^{sync}$};
   \node[] (c8) at (8,1.5) {$k_8$};

  \path[->]
    (c0s) edge node {} (c1d)
    (c1d) edge node [above] {$k_2$} (c2d)
          edge node [above] {$k_3$} (c3d)
          edge [bend left=80] node [above] {$k_{1;4}^{sync}$} (c2d)
    (c2d) edge node [above] {$k_4$} (c4d)
    (c4d) edge node [above] {$k_6$} (c6d)
    (c6d) edge node [above] {} (c8d)
    (c3d) edge node [above] {$k_0$} (c6d)

@@

   %reduction c0 c1
   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,2.5) {$k_0$};
   \node[] (star1) at (1.5,2.5) {$k_{0;1}^{sync}$};
   \node[] (c1) at (2,1.5) {$k_1$};

   %reduction c7 c8
   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
   \node[] (c7) at (6.5,2.5) {$k_7$};
   \node[] (star8) at (7.5,2.5) {$k_{7;8}^{sync}$};
   \node[] (c8) at (8,1.5) {$k_8$};

   %reduction c3 c5
   \node[] (s5) at (4.5,-2) {$\mathcal{S}$};
   \node[] (c3) at (4,-1) {$k_3$};
   \node[] (c5) at (5,-1) {$k_0$};

   %reduction c2 *4
   \node[] (p0) at (3.5,1) {$\mathcal{P}$};
   \node[] (c2) at (3,2) {$k_2$};
   \node[] (star4) at (4,2) {$k_{1;4}^{sync}$};

  \path[->]
    (c0s) edge node {} (c1d)
    (c1d) edge node [above] {} (c2d)
    (c2d) edge node [above] {$k_4$} (c4d)
    (c4d) edge node [above] {$k_6$} (c6d)
    (c6d) edge node [above] {} (c8d)
    (c1d) edge [bend right] node [above] {} (c6d)

@@

    %reduction c2 *4
    (p0) edge node {} (c2)
         edge node {} (star4);
  \end{tikzpicture}
}
\end{center}
\caption{Sequence and parallel rules applications}
\label{fig:SP}
\end{figure}

\begin{figure}[ht!]
\begin{center}
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

   %reduction c0 c1
   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,2.5) {$k_0$};
   \node[] (star1) at (1.5,2.5) {$k_{0;1}^{sync}$};
   \node[] (c1) at (2,1.5) {$k_1$};

   %reduction c7 c8
   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
   \node[] (c7) at (6.5,2.5) {$k_7$};
   \node[] (star8) at (7.5,2.5) {$k_{7;8}^{sync}$};
   \node[] (c8) at (8,1.5) {$k_8$};

   %reduction c3 c5
   \node[] (s5) at (4.5,-2) {$\mathcal{S}$};
   \node[] (c3) at (4,-1) {$k_3$};
   \node[] (c5) at (5,-1) {$k_0$};

   %reduction c2 c6
   \node[] (s6) at (4.5,1) {$\mathcal{S}$};
   \node[] (p0) at (3.5,2) {$\mathcal{P}$};
   \node[] (c2) at (3,3) {$k_2$};
   \node[] (star4) at (4,3) {$k_{1;4}^{sync}$};
   \node[] (s7) at (5.5,2) {$\mathcal{S}$};
   \node[] (c4) at (5,3) {$k_4$};
   \node[] (c6) at (6,3) {$k_6$};

@@

   %reduction c0 c1
   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,2.5) {$k_0$};
   \node[] (star1) at (1.5,2.5) {$k_{0;1}^{sync}$};
   \node[] (c1) at (2,1.5) {$k_1$};

   %reduction c7 c8
   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
   \node[] (c7) at (6.5,2.5) {$k_7$};
   \node[] (star8) at (7.5,2.5) {$k_{7;8}^{sync}$};
   \node[] (c8) at (8,1.5) {$k_8$};

   \node[] (p1) at (4.5,0.5) {$\mathcal{P}$};
   %reduction c3 c5
   \node[] (s5) at (5.5,1.5) {$\mathcal{S}$};
   \node[] (c3) at (5,2.5) {$k_3$};
   \node[] (c5) at (6,2.5) {$k_0$};
   %reduction c2 c6
   \node[] (s6) at (3.5,1.5) {$\mathcal{S}$};
   \node[] (p0) at (2.8,2.5) {$\mathcal{P}$};
   \node[] (c2) at (2.3,3.5) {$k_2$};
   \node[] (star4) at (3.3,3.5) {$k_{1;4}^{sync}$};
   \node[] (s7) at (4.2,2.5) {$\mathcal{S}$};
   \node[] (c4) at (3.8,3.5) {$k_4$};
   \node[] (c6) at (4.8,3.5) {$k_6$};

@@

   %reduction c0 c1
   \node[] (s0) at (1.5,2.5) {$\mathcal{S}$};
   \node[] (s1) at (1,3.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,4.5) {$k_0$};
   \node[] (star1) at (1.5,4.5) {$k_{0;1}^{sync}$};
   \node[] (c1) at (2,3.5) {$k_1$};

   %reduction c7 c8
   \node[] (s2) at (7.5,1.5) {$\mathcal{S}$};
   \node[] (s3) at (7,2.5) {$\mathcal{S}$};
   \node[] (c7) at (6.5,3.5) {$k_7$};
   \node[] (star8) at (7.5,3.5) {$k_{7;8}^{sync}$};
   \node[] (c8) at (8,2.5) {$k_8$};

   \node[] (p1) at (4.5,2.5) {$\mathcal{P}$};
   %reduction c3 c5
   \node[] (s5) at (5.5,3.5) {$\mathcal{S}$};
   \node[] (c3) at (5,4.5) {$k_3$};
   \node[] (c5) at (6,4.5) {$k_0$};
   %reduction c2 c6
   \node[] (s6) at (3.5,3.5) {$\mathcal{S}$};
   \node[] (p0) at (2.8,4.5) {$\mathcal{P}$};
   \node[] (c2) at (2.3,5.5) {$k_2$};
   \node[] (star4) at (3.3,5.5) {$k_{1;4}^{sync}$};
   \node[] (s7) at (4.2,4.5) {$\mathcal{S}$};
   \node[] (c4) at (3.8,5.5) {$k_4$};
   \node[] (c6) at (4.8,5.5) {$k_6$};

@@

    (p1) edge node {} (s6)
         edge node {} (s5)
    %reduction c3 c5
    (s5) edge node {} (c3)
         edge node {} (c5)
    %reduction c2 c6
    (s6) edge node {} (p0)
         edge node {} (s7)
    %reduction c2 *4
    (p0) edge node {} (c2)
         edge node {} (star4)
    (s7) edge node {} (c4)
         edge node {} (c6);
  \end{tikzpicture}
}
\end{center}
\caption{Rest of sequence and parallel rules.}
\label{fig:example}
\end{figure}

### 2016-01-29 ARTICLES/IJPP/article/formalism.tex
article IJPP Helene 26/01/2016
article IJPP corrections Helene 29/01/2016

$comp$ represents the actual numerical expression which is computed by a kernel. We can dissociate three types of kernel computations.

\noindent \textbf{Definitions}
%\begin{mydef}
\begin{itemize}
\item We denote by $identity$ the identity function $x=x$.
\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{stencil kernel} $\iff \exists (r,n) \in R$ such that $n \neq identity$.
\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{local kernel} $\iff \forall (r,n) \in R$, $n = identity$.
\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{reduction kernel} $\iff w$ is a scalar, $R=\{(r,n)\}$, and if $n=entity(r)$.
\end{itemize}

@@

\medskip
\noindent \textbf{Definitions}
\begin{itemize}
\item The set of $n$ ordered computation kernels of a numerical simulation is denoted $\Gamma = [k_i]_{0 \leq i \leq n-1}$, such that $\forall k_i,k_j \in \Gamma$, if $i \leq j$, then $k_i$ is computed before $k_j$.
\item Finally, a \textit{multi-stencil program} is defined by the septuplet 
\begin{equation*}
\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta, \mathcal{S},\Gamma)
\end{equation*}
\end{itemize}

>>>>

$comp$ represents the actual numerical expression which is computed by a kernel. We can dissociate four types of kernel computations.

\noindent \textbf{Definitions}
%\begin{mydef}
\begin{itemize}
\item We denote by $identity$ the identity function $x=x$.
\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{stencil kernel} $\iff \exists (r,n) \in R$ such that $n \neq identity$ and if $\exists (r,n)$ with $r=w$ then $n=identity$.
\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{boundary kernel} $\iff \exists (r,n) \in R$ such that $r=w$ and such that $n \neq identity$.
\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{local kernel} $\iff \forall (r,n) \in R$, $n = identity$.
\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{reduction kernel} $\iff w$ is a scalar, $R=\{(r,n)\}$, and if $n=entity(r)$.
\end{itemize}

In our formalism, it is forbidden for a stencil kernel to find $(r,n) \in R$ for which $r=w$ and $n \neq identity$. Actually, if such a property is possible implicit numerical schemes are needed into the simulation. Such a scheme is not a stencil and is over the scope of this paper. The problem does not happened for local kernels as $\forall (r,n) \in R$, $n = identity$. Computations for which this property is authorized are boundary kernels only. This type of kernel is a particular case as it commputes boundary conditions, which represents what should happened outside the space domain but still impact stencil kernels at the next time step.

@@

\medskip
\noindent \textbf{Definitions}
\begin{itemize}
\item The set of $n$ ordered computation kernels of a numerical simulation is denoted $\Gamma = [k_i]_{0 \leq i \leq n-1}$, such that $\forall k_i,k_j \in \Gamma$, if $i \leq j$, then $k_i$ is computed before $k_j$.
\item Finally, a \textit{multi-stencil program} is defined by the octuplet 
\begin{equation*}
\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\mathcal{N},\Delta, \mathcal{S},\Gamma)
\end{equation*}
\end{itemize}
